---
title: "Shooting Incidents Reported by NYPD"
author: "Jungho Kwon"
date: "2023-06-08"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Let's analyze the shooting incidents reported by the NYPD from January 2006 to December 2022.

The data is obtained from the City of New York ([opendata.cityofnewyork.us](https://opendata.cityofnewyork.us/)).

\vspace{2cm}

## Import and clean the data
```{r read_data, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)

url_in <- "https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD"
nypd_data = read_csv(url_in)
nypd_data <- nypd_data %>%
  select(-c(Latitude, Longitude, X_COORD_CD, Y_COORD_CD, Lon_Lat, JURISDICTION_CODE )) %>%
  mutate(OCCUR_DATE = mdy(OCCUR_DATE)) %>%
  replace_na(list(LOCATION_DESC = 'NONE')) %>%
  replace_na(list(PERP_AGE_GROUP = 'UNKNOWN', PERP_SEX='UNKNOWN', PERP_RACE='UNKNOWN'))
```

## Group by month
```{r group_by_month, message=FALSE, warning=FALSE}
by_month <- nypd_data %>% 
  group_by(month = floor_date(OCCUR_DATE, "month")) %>% 
  distinct(month, INCIDENT_KEY) %>% 
  select(month, INCIDENT_KEY) %>% 
  tally()
```

## Plot the number of incidents over time
```{r plot_group_by_month, message=FALSE, warning=FALSE}
by_month %>% ggplot(aes(x = month, y = n)) +
  geom_point(aes(color="Number of incidents in a month")) +
  theme(legend.position = "bottom") +
  xlab("Time") + ylab("Number of incidents") + 
  labs(title = "Number of shooting incidents reported by NYPD for each month")
```

## Create a linear model and prediction
```{r plot_linear_model, message=FALSE, warning=FALSE}
mod <- lm(n ~ month, data = by_month)

by_month_pred <- by_month %>% 
  mutate(pred = predict(mod))

by_month_pred %>%
  ggplot() +
  geom_point(aes(x = month, y = n), color="red") +
  geom_line(aes(x = month, y = pred), color="black") +
  theme(legend.position = "bottom") +
  xlab("Time") + ylab("Number of incidents") + 
  labs(title = "Number of shooting incidents reported by NYPD for each month")
```

The black line on the graph is a prediction calculated from the linear model of time and the number of shooting incidents.

\newpage

## Group by each borough
```{r by_boro, message=FALSE, warning=FALSE}
by_month_BRONX <- nypd_data %>%
  filter(BORO == "BRONX") %>%
  group_by(month = floor_date(OCCUR_DATE, "month"), BORO) %>% 
  distinct(month, INCIDENT_KEY, BORO) %>% 
  select(month, INCIDENT_KEY, BORO) %>% 
  tally()
  
by_month_BROOKLYN <- nypd_data %>%
  filter(BORO == "BROOKLYN") %>%
  group_by(month = floor_date(OCCUR_DATE, "month"), BORO) %>% 
  distinct(month, INCIDENT_KEY, BORO) %>% 
  select(month, INCIDENT_KEY, BORO) %>% 
  tally()

by_month_MANHATTAN <- nypd_data %>%
  filter(BORO == "MANHATTAN") %>%
  group_by(month = floor_date(OCCUR_DATE, "month"), BORO) %>% 
  distinct(month, INCIDENT_KEY, BORO) %>% 
  select(month, INCIDENT_KEY, BORO) %>% 
  tally()

by_month_QUEENS <- nypd_data %>%
  filter(BORO == "QUEENS") %>%
  group_by(month = floor_date(OCCUR_DATE, "month"), BORO) %>% 
  distinct(month, INCIDENT_KEY, BORO) %>% 
  select(month, INCIDENT_KEY, BORO) %>% 
  tally()

by_month_STATEN_ISLAND <- nypd_data %>%
  filter(BORO == "STATEN ISLAND") %>%
  group_by(month = floor_date(OCCUR_DATE, "month"), BORO) %>% 
  distinct(month, INCIDENT_KEY, BORO) %>% 
  select(month, INCIDENT_KEY, BORO) %>% 
  tally()
```

## Plot the number of incidents for each borough over time
```{r plot_by_boro, message=FALSE, warning=FALSE}
ggplot() +
  geom_point(data = by_month_BRONX, aes(x = month, y = n, color = "Bronx")) +
  geom_point(data = by_month_BROOKLYN, aes(x = month, y = n, color = "Brooklyn")) +
  geom_point(data = by_month_MANHATTAN, aes(x = month, y = n, color = "Manhattan")) +
  geom_point(data = by_month_QUEENS, aes(x = month, y = n, color = "Queens")) +
  geom_point(data = by_month_STATEN_ISLAND, aes(x = month, y = n, color = "Staten Island")) +
  scale_y_log10() +
  theme(legend.position = "bottom") +
  xlab("Time") + ylab("Number of incidents") + 
  labs(title = "Number of shooting incidents reported by NYPD for each borough")
```


# Conclusion

Looking at the linear model, it seemed as if the number of shooting incidents was decreasing over time (in
New York from 2006 to 2020). In 2020, there is an increase in the number of shooting incidents.

Some questions come to mind

* Did COVID and the initial lockdowns play a role in the increase of shooting incidents in 2020? For
me, the year 2020 has been associated with the COVID outbreak and initial lockdowns. I need to be
careful in asking such questions, as there is not enough data / variables.
* What is the population density in each boroughs of New York?
* How does each precinct report on the incidents?

I have a bias, where I think people who live in cities tend to be more uptight (thus, more conflicts) compared
to those who live out side the city (Staten Island). This could make me think incorrectly that this is the
reason why there are more shooting incidents reported near the city center (Manhattan, Queens, Bronx,
Brooklyn) vs Staten Island. However, we would need to look at the population density before comparing
the number of incidents between boroughs.

\newpage

## Session Info for debugging
```{r session_info, echo=TRUE}
sessionInfo()
```
